---
title: "College softball"
author: "Emma Caetano"
---

```{r}
#| label: setup
#| include: false

library(tidyverse)
library(reactable)
library(plotly)

# Read the 2026 college softball WAR data
war_data <- read_csv("pennant-war-2026.csv", show_col_types = FALSE)
```

### Batting WAR vs. Pitching WAR

```{r}
#| label: war-scatter-plot
#| echo: false

# Interactive scatter plot showing offense vs pitching balance
p_scatter <- war_data |>
  ggplot(aes(
    x = war_bat,
    y = war_pit,
    text = paste0(
      "<b>", team, "</b> (", conference, ")<br>",
      "Rank: ", war_rank, "<br>",
      "Total WAR: ", round(war_total, 2), "<br>",
      "Batting WAR: ", round(war_bat, 2), "<br>",
      "Pitching WAR: ", round(war_pit, 2)
    )
  )) +
  geom_point(aes(color = conference), alpha = 0.7, size = 2) +
  theme_minimal() +
  labs(
    x = "Batting WAR",
    y = "Pitching WAR"
  ) +
  theme(legend.position = "none")

ggplotly(p_scatter, tooltip = "text")
```

Softball hitting WAR (Wins Above Replacement) is an advanced sabermetric statistic that estimates how many more wins a specific batter contributes to their team compared to a free, readily available "replacement-level" player. It goes beyond simple metrics like batting average or home runs by looking at a player's complete batting contribution (using measures like weighted on-base average or linear weights). Softball pitching WAR is an advanced sabermetric statistic that estimates the number of wins a pitcher adds to her team compared to a hypothetical "replacement-level" pitcher (a readily available backup or average reserve player).

