---
title: "Leaderboard"
author: "Emma Caetano"
---

Softball hitting WAR (Wins Above Replacement) is an advanced sabermetric statistic that estimates how many more wins a specific batter contributes to their team compared to a free, readily available "replacement-level" player. It goes beyond simple metrics like batting average or home runs by looking at a player's complete batting contribution (using measures like weighted on-base average or linear weights). Softball pitching WAR is an advanced sabermetric statistic that estimates the number of wins a pitcher adds to her team compared to a hypothetical "replacement-level" pitcher (a readily available backup or average reserve player).

```{r}
#| label: setup
#| include: false

library(tidyverse)
library(reactable)

war_data <- read_csv("pennant-war-2026.csv", show_col_types = FALSE)
```

```{r}
#| label: war-table
#| echo: false

war_data |>
  select(
    Rank = war_rank,
    Team = team,
    Conference = conference,
    `Batting WAR` = war_bat,
    `Pitching WAR` = war_pit,
    `Total WAR` = war_total
  ) |>
  reactable(
    searchable = TRUE,
    filterable = TRUE,
    striped = TRUE,
    defaultPageSize = 10,
    columns = list(
      Team = colDef(filterable = TRUE),
      Conference = colDef(filterable = TRUE),
      `Batting WAR` = colDef(format = colFormat(digits = 2)),
      `Pitching WAR` = colDef(format = colFormat(digits = 2)),
      `Total WAR` = colDef(format = colFormat(digits = 2))
    )
  )
```
